---
title: "Combination Permutation Recap"
format: revealjs
---

## Agenda

-   Announcements
-   Counting, Combination, Permutation Practice
-   Combination - $n$ choose $r$ refresher

## Concept Question 1 - Combinations

How many different outfits can you create with 3 pairs of pants with 4 shirts and 2 pairs of shoes?

```{r}
countdown::countdown(3, top = 2.5)
```

## Permutation Review

-   In permutations, order matters.

Number of permutations of $n$ different objects

$$
n \times (n-1) \times \cdots 2 \times 1 = n!
$$

-   How many ways could you line up everyone (9 people) in the class?

## Concept Question: Permutation of r groups of n objects

Among $n$ objects, if $n_1$ are alike, $n_2$ are alike, $\ldots$, $n_r$ are alike, then there are

$$
\frac{n!}{n_1! n_2! \cdots n_r!}
$$

different permutations.

How many different permutations for the word *acclimate*?

Working with a partner - write your solution on the board.

```{r}
countdown::countdown(3, top = 2.5)
```

------------------------------------------------------------------------

## Combinations recap - order doesn't matter practice

-   If we have $n$ items and want to select $r$ of them,

$$
\binom{n}{r} = \frac{n!}{(n-r)!r!}
$$

The number ${n \choose r}$ is pronounced as $n$ choose $r$, it is the number of ways to pick $r$ objects from a set of $n$ distinct objects.
