---
title: "Combination Permutation Differences"
format: revealjs
---

## Agenda

-   Announcements
-   Combinations vs. Permutations
-   Combinations vs. Permuatations Definitions
-   Use Cases
-   Formulas
-   Solutions to problems

## Combinations and Permutations

- Fundamental concepts in probability and combinatorics* that deal with the **selection** and **arrangement** of objects.

*area of mathematics primarily concerned with counting things.

## Definitions

- **Permutations** refer to the different ways in which a set of objects can be arranged, where the order does matter.

- **Combinations** refer to the different ways in which a set of objects can be selected, where the order *does not* matter.

## Order Matters

- **Permutations**: The order of selection is important. For example, if you are arranging a presentation, the sequence in which the presentation is given matters (Introduction, Methods, Conclusion is different from Conclusion, Methods, Introduction).

- **Combinations**: The order of selection is not important.For example, if you are forming a committee of 3 members out of 5 people, the sequence in which you pick the members does not matter (Alice, Julio, Saliha is the same as Saliha, Alice, Julio).

## Use Cases

- **Permutations** are used when we need to determine the number of possible arrangements of a set of items where their order matters (e.g., seating arrangements, word formation).

- **Combinations** are used when we need to determine the number of ways to select items from a set where their order does not matter (e.g., sandwich fillings, selecting a team).

**What are some other use cases you can think of?**

```{r}
countdown::countdown(3, top = 3.5)
```

## Formulas Review - Permutations

Number of permutations of $n$ different objects

$$
n \times (n-1) \times \cdots 2 \times 1 = n!
$$

**Variations**: r groups of n objects

Among $n$ objects, if $n_1$ are alike, $n_2$ are alike, $\ldots$, $n_r$ are alike, then there are

$$
\frac{n!}{n_1! n_2! \cdots n_r!}
$$

different permutations.

How many different permutations for the word *epitomize*?


## Formula Review - **Counting combinations**

Suppose an experiment consists $r$ different outcomes, with the $i$-th outcome having $n_i$ possibilities, then together there are

$$
n_1 \times n_2 \times \cdots \times n_r = \prod_{i=1}^{r} n_i
$$

How many license combinations could be made if the license must consist to 3 letters and 3 numbers?

```{r}
countdown::countdown(3, top = 3.5)
```

## Formula Review - **Combinations**

- If we have $n$ items and want to select $r$ of them,


$$
\binom{n}{r} = \frac{n!}{(n-r)!r!}
$$

The number ${n \choose r}$ is pronounced as $n$ choose $r$, it is the number of ways to pick
$r$ objects from a set of $n$ distinct objects.

*Here, $r!$ accounts for the fact that the order in which the $r$ objects are chosen does not matter.


## Summary

- **Permutations** = Selection with arrangement (order matters).

- **Combinations** = Selection without arrangement (order does not matter).

Understanding whether the order is important or not. is key to distinguishing between the two.

## Question 2: Forming a Faculty-Student Committee

A college is forming a new committee to improve campus facilities, which will consist of 5 members: 3 faculty members and 2 students. If there are 8 faculty members and 7 students available, how many ways can the committee be formed if it must include exactly 3 faculty members and 2 students?

## Question 2 Solution: Step 1

This is a combination problem because the order in which the faculty members and students are chosen does not matter.

We need to select: 

- 3 faculty members from 8, 
- 2 students from 7.

## Choose 3 faculty out of 8

First, calculate the number of ways to choose 3 faculty members out of 8:

$$
\binom{n}{r} = \frac{n!}{(n-r)!r!}
$$

$$
\binom{8}{3} = \frac{8!}{3! \cdot (8 - 3)!} = \frac{8!}{3! \cdot 5!}
$$

$$
\binom{8}{3} = \frac{8 \times 7 \times 6}{3 \times 2 \times 1} = \frac{336}{6} = 56
$$
## Choose 2 students out of 7
Next, calculate the number of ways to choose 2 students out of 7:

$$
\binom{7}{2} = \frac{7!}{2! \cdot (7 - 2)!} = \frac{7!}{2! \cdot 5!}
$$

$$
\binom{7}{2} = \frac{7 \times 6}{2 \times 1} = \frac{42}{2} = 21
$$
## Question 2 Solution: Step 2

Now, multiply the number of ways to choose faculty members and students to get the total number of ways to form the committee:



$$
\text{Total ways} = \binom{8}{3} \times \binom{7}{2} = 56 \times 21 = 1176
$$

There are 1,176 ways to form the committee.


