---
title: "Introduction to Probability Recap"
format: revealjs
---

## Agenda

-   Announcements
-   Review of ideas in reading
-   Probability refresher
-   Concept Questions
-   Problem Set 1: Introducing Probability

## Review of syllabus questions

-   Grading Contract Demo
-   Question Forum Demo

# Concept review: Rules

## Rules of probability {.incremental}

Let $\Omega$ be the outcome space, and let $P(A)$ denote the probability of the event $A$. Then we have:

. . .

1.  $P(A) \ge 0$

. . .

2.  $P(\Omega) = 1$

. . .

3.  If $A$ and $B$ are mutually exclusive ($A \cap B = \{\}$), then $P(A \cup B) = P(A) + P(B)$

::: notes
Concepts to review: (KEEP THIS BRIEF, OR INCORPORATE INTO CQ's) - The first two rules of probability - unions and intersections - mutually exclusive events and the addition rule - good idea to draw Venn diagrams here - Use rule 3 to write down the complement rule, and show what A\^C means
:::

# Probability refresher

```{r}
countdown::countdown(5, top = 0)
```

::: notes
Give them five minutes, can use kahoot music
:::

# Concept Question 1

## The Linda Problem {.scrollable}

```{r}
countdown::countdown(1, top = 0)
```

The Linda problem is from a very famous experiment conducted by Daniel Kahneman and Amos Tversky in 1983.

Linda is thirty-one years old, single, outspoken, and very bright. She majored in philosophy. As a student, she was deeply concerned with issues of discrimination and social justice, and also participated in antinuclear demonstrations.

::: poll
Which alternative is more probable?
:::

(a) Linda is a bank teller.

(b) Linda is a bank teller and is active in the feminist movement.

::: notes
Correct answer: (a) Depending on the response, you can discuss how even though (b) is clearly contained in (a) and therefore has lower probability, an overwhelming majority of their respondents ranked (b) as more likely. "About 85% to 90% of undergraduates at several major universities chose the second option, contrary to logic", and talk about why this is so. Probability can be tricky and counter-intuitive. If they do well, congratulate them and say that they are among the rare people who understand that $P(A \textbf{ and } B)$ *must* be lower than $P(A)$.
:::

Kahneman, Daniel. Thinking, Fast and Slow (p. 158). Farrar, Straus and Giroux.

# Problem Set

-   You may work individually, in pairs or in threes for this assignment.
-   Ask your neighbor a question first, and then put a sticky note on your desk if you want me to stop by.

```{r}
countdown::countdown(30, top = 0)
```

# Intro to web-r

-   [Brief Intro R and Sampling Functions](https://laurielbaker.github.io/coa-probability-webr-labs/lab-intro-r-and-sampling-functions.html)

```{r}
countdown::countdown(3, top = 0)
```
