---
title: "Combination Permutation Recap"
format: revealjs
---

## Agenda

-   Announcements
-   Web-R Review
-   Counting, Combination, Permutation Refresher
-   Concept Questions
-   Practice Problems: Combination and Permutation

## Web-R Review

-   [Web-r exercises]()

## The Basic Rule of Counting Review

Suppose an experiment consists $r$ different outcomes, with the $i$-th outcome having $n_i$ possibilities, then together there are

$$
n_1 \times n_2 \times \cdots \times n_r = \prod_{i=1}^{r} n_i
$$

In our Beano game, we have two six-sided die. Following this rule we can figure out the number of possible combinations as

$$
6 \times 6 = 36
$$

## Concept Question 1

A sandwich shop has 3 bread options, 4 spread options, and 4 fillings. How many different sandwiches can you create?

## Permutations Review

-   Using the cards given to each pair. List and count the number of ways that the cards could be arranged and report back that number.

```{r}
countdown::countdown(2, top = 2.5)
```

## Permutation Review

-   In permutations, order matters.

For two cards we have $2!$ options:

$$
2 \times 1
$$

For three cards we have $3!$ options:

$$
3 \times 2 \times 1
$$

Number of permutations of $n$ different objects

$$
n \times (n-1) \times \cdots 2 \times 1 = n!
$$

## Permutations in R

In R, `factorial()` is a function

```{r echo = TRUE}
factorial(2)
```

```{r echo = TRUE}
factorial(3)
```

```{r echo = TRUE}
factorial(100)
```

## Recap Permutations

How many permutations in the word Facebook?

-   $\frac{8!}{2!}$

|     |     |     |     |     |       |       |     |
|-----|-----|-----|-----|-----|-------|-------|-----|
| c   | e   | F   | a   | b   | o$_1$ | o$_2$ | k   |
| c   | e   | F   | a   | b   | o$_2$ | o$_1$ | k   |

-   Why? What does each part represent?

```{r}
countdown::countdown(2, top = 2.5)
```

## Permutation of r groups of n objects

Among $n$ objects, if $n_1$ are alike, $n_2$ are alike, $\ldots$, $n_r$ are alike, then there are

$$
\frac{n!}{n_1! n_2! \cdots n_r!}
$$

different permutations.

How many different permutations for the word *bananagrams*?

Working with a partner - write your solution on the board.

```{r}
countdown::countdown(3, top = 2.5)
```
